
# Maps of collected biodiversity by Colombian institutions

In a previous [post](http://rpubs.com/dlizcano/6337) I got the georeferenced data set of biodiversity from [SIB](http://data.sibcolombia.net/conjuntos/) Colombia using [IAvH code](http://code.google.com/p/descarga-valida-gbif-sib/). The data set is composet of 127 tables corresponding to the [GBIF](http://www.gbif.org/) grid. 

Some tables still need some additional work to fix extra spaces and inconsistent characters. But after some work, I put together all the tables. When plotting the data, you can see where Colombian researchers have mostly collected biodiversity.

This time I want to see the use of ggplot2 and ggmaps to discover the contribution of each institution.

## code chunks

The data set is 70 megas aprox.  

```{r makemaps, echo=TRUE, comment=NA, tidy=TRUE}

require(ggmap)
require(raster)
#### load data set 252.944 records
bigtable<-read.csv(file="data/sib_bigtable.csv",header=T,encoding = "windows-1252")

#get poligon Colombia
co<-getData("GADM",country="CO",level=1,download=TRUE)
col_depto <- fortify(co,region="NAME_1") # make compatible to ggplot2

locat=as.vector(bbox(co))
ncmap = get_map(location=locat,source="stamen",maptype="toner",zoom=6)
#ggmap(ncmap) not nice

mapbase<- ggplot(col_depto, aes(long,lat,group=group)) + geom_polygon(fill="grey60") + coord_equal() +
  geom_path(color="grey") 

mapbase2<- ggplot(col_depto, aes(long,lat,group=group)) + geom_polygon(fill="White ") + coord_equal() +
  geom_path(color="grey") 

map1<-mapbase2 + geom_point(aes(x = lon, y = lat, group = TRUE), 
                    data = bigtable, size = 1.5,
                    alpha = 1/20) +
                    theme(legend.position = "right") +
                    guides(guide_legend((title = NULL)))

map2<-mapbase + geom_point(aes(x = lon, y = lat, group = FALSE), size=1, 
                    data = bigtable,alpha=I(0.25),colour="steelblue") + 
                    stat_binhex(aes(x = lon, y = lat, group = FALSE),
                    size = .5, binwidth = c(.5,.5), 
                    alpha = 2/4,data =bigtable)

map3<-mapbase + geom_point(aes(x = lon, y = lat, group = TRUE,
                     colour = factor(institution)), 
                     data = bigtable, size = 2,
                     alpha=1/2) + 
                     theme(legend.position = "right") 

map4<-mapbase + geom_point(aes(x = lon, y = lat, group = FALSE), 
                    size=0.5, 
                    data = bigtable,colour="red") + 
                    facet_wrap(~ institution, ncol = 6)
  

```

## The first map. All

A map showing all collection points. The points are transparent. So darker means more points in that location. 
```{r map1, comment=NA, tidy=TRUE}
map1
```

## The second map. Collected Where?

A map showing hexagonal bins with more collection points. Interesting: no hexagons means no colecction in that place.
```{r map2, comment=NA, tidy=TRUE}
map2
```

## The third map by institution

A map showing all the points. Colors by institution.
```{r map3, comment=NA, fig.width=8, tidy=TRUE}
map3
```

## The last map wraped by institution

This is slow, but worth.... meanwhile take a coffee.
```{r map4, comment=NA, fig.height=7,fig.width=8,fig.align='center',tidy=TRUE}
map4
```

### It is nice to see how the Herbario Nacional Colombiano (COL) is the most extensive collection in Colombia. Now I have to figure which colection or institution is 8200001422-01. 

After some coments from [@OigaMen](https://twitter.com/OigaMen) and [@Danipilze](https://twitter.com/Danipilze) now I know that the code 8200001422-01 is part of the IAvH collection. So for the next post I am going to fix that.

I had fun learning how to publish blog posts [from R + knitr to WordPress](http://yihui.name/knitr/demo/wordpress/). Next posts will even more interesting, I am sure.  


